The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging with the Owens Valley Long Wavelength Array

### 2017-06-28 paper.tex
expanded error analysis section

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figures/antenna-layout}
    \caption{
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 open circles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to 1.5 km. The 5 crosses are
        antennas equipped with noise-switched front ends.
    }
    \label{fig:antenna-layout}
\end{figure}

The flux scale is tied to the work of \citet{2016arXiv160905940P}.

Without removing bright sources from the data, sidelobes from bright sources will dominate the
variance in the image.  At 74 MHz Cyg A is a 15,000 Jy source \citep{2016arXiv160905940P}. A
conservative estimate for the confusion limit at 74 MHz with a 10 arcminute beam is 500 mJy
\citep{2012RaSc...47.0K04L}. Therefore we require that Cyg A's sidelobes be at most of $-$45 dB down
from its peak flux to prevent Cyg A's sidelobes from dominating the variance in the image.
\todo{Measure Cyg A's sidelobe levels if it is not subtracted}

\subsubsection{Sparking Power Lines}

The Owens Valley is an important source of water and power for the city of Los Angeles.
Unfortunately this means that high voltage power lines run along the valley to the west of the OVRO-
LWA. Some of these power line poles have faulty insulators that arc and produce pulsed, broadband
RFI. Because these poles exist in the near-field of the array, we have been able to localize some of
them by using the curvature of the incoming wavefront. Efforts are currently underway to work with
the utility pole owners to have these insulators replaced.

In the meantime it is possible to suppress their contamination in the dataset. The contribution of
these RFI sources to the visibilities can be plainly seen by averaging $>$24 hours of data with the
phase center set to zenith. In this way, true sky components are smeared along tracks of constant
declination while terrestrial sources (ie. the arcing power lines) are not smeared. For similar
reasons, when constructing a map of the sky from $>$24 hours of data these sparking power lines will
produce artifacts in the map that are smeared along tracks of constant declination. However while
the amplitude of the RFI will fluctuate with time, its phase will be constant. Therefore if you can
use isolate a set of model visibilities for the RFI, it can be removed from each integration by
scaling and subtracting these model visibilities from the measured visibilities.

Obtaining a model for the RFI is complicated by the fact that the contaminating sources are at
extremely low elevations where the antenna response is maximally inhomogeneous. Therefore it is not
enough to know the physical location of the faulty insulator generating the RFI. In addition you
must also know the response of each antenna (amplitude and phase) in the appropriate direction. This
motivates the use of peeling, which allows the antenna response to be a free parameter.

Therefore model visibilities for the RFI can be obtained by peeling the sources after smearing the
visibilities over $>$24 hours. These model visibilities are then scaled and subtracted from each
integration in the data set.

\todo{show an example smeared image before and after peeling}

%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cccc}
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw04-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw06-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw08-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw10-small} \\
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw12-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw14-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw16-small} &
%        \includegraphics[width=0.22\textwidth]{figures/smeared-snapshots/spw18-small} \\
%    \end{tabular}
%    \caption{
%        Dirty snapshot images after smearing the visiblities over the entire 100 hour observation
%        window.  The two bright source tracks correspond to Cas A (top track) and Cyg A (bottom
%        track) smeared along a ring of constant declination. The sources dotting the horizon are
%        faulty power line insulators.
%    }
%\end{figure*}

\section{Imaging}

@@

\section{Errors}

\subsection{Terrestrial Interference}

When writing down equation~\ref{eq:basic-imaging}, it is implicitly assumed that the correlated
voltage fluctuations measured between pairs of antennas are generated by astronomical sources of
radio emission. In practice, this implicit assumption can be violated. For instance a low-frequency
interferometer located in the vicinity of an arcing power line will see an additional contribution
from the radio-frequency interference (RFI) generated by the arcing process. Similarly common-mode
pickup along the analog signal path of the interferometer may generate an additional spurious
contribution to the measured visibilities. While the amplitude and phase of these contaminating
signals may fluctuate with time, they do not sweep across the sky at the sidereal rate
characteristic of astronomical sources.

@@

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = 0$.  In
practice this is unlikely. In general the contamination is proportional to the overall amplitude of
the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/\|\b v_\text{terrestrial}\|$).

These astronomical sources are anchored to the celestial sphere and therefore rotate
across the sky over the course of a sidereal day. Terrestrial source of interference are not
anchored to the celestial sphere.

This characteristic motion of astronomical sources
is a fundamental component to $m$-mode analysis imaging

radio emission is astronomical. Astronomical sources are anchored to the celestial sphere and
therefore rotate across the sky over the course of a sidereal day. However, in a practice

$m$-mode analysis imaging implicitly assumes that the correlation between any two pairs of antennas

However it is instructive to analyze the contaminating effect the RFI will have
on $m$-mode analysis maps of the sky. The correlation observed on a given
baseline and frequency channel for a single source of RFI is
\begin{equation}
    V(t) = I(t) B^\textrm{near-field}
\end{equation}

However we need a strategy to mitigate the RFI before imaging. A constant source attached to the
horizon has a fringe rate of zero. The only possible astronomical origins for a source with zero
fringe rate are:

\begin{enumerate}
    \item a point source located at the north or south celestial poles, and
    \item a ring of constant declination.
\end{enumerate}

Therefore while snapshot images will show the RFI source on the horizon, extended integrations will
show the RFI as rings of constant declination. We can take advantage of this discrepancy to separate
astronomical emission from terrestrial RFI.

@@

\begin{itemize}
    \item declination dependent flux scale errors
    \item failure to reconcile observations from two separate times
\end{itemize}

\subsection{Ionospheric Scintillation}

@@

This will tend to scatter power between $m$-modes.

Simulate images of point sources with this

\subsection{Ionospheric Refraction}

>>>>

\section{All-Sky Imaging}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figures/antenna-layout/antenna-layout}
    \caption{
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 open circles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to 1.5 km. The 5 crosses are
        antennas equipped with noise-switched front ends.
    }
    \label{fig:antenna-layout}
\end{figure}

The flux scale is tied to the work of \citet{1977A&A....61...99B, 2012MNRAS.423L..30S,
2017ApJS..230....7P}.

Without removing bright sources from the data, sidelobes from bright sources will dominate the
variance in the image.  At 74 MHz Cyg A is a 15,000 Jy source \citep{2017ApJS..230....7P}. A
conservative estimate for the confusion limit at 74 MHz with a 10 arcminute beam is 500 mJy
\citep{2012RaSc...47.0K04L}. Therefore we require that Cyg A's sidelobes be at most of $-$45 dB down
from its peak flux to prevent Cyg A's sidelobes from dominating the variance in the image.
\todo{Measure Cyg A's sidelobe levels if it is not subtracted}

\section{Error Analysis}

\subsection{Terrestrial Interference and Pickup}

\begin{figure*}[ht]
    \centering
    \begin{tabular}{cccc}
        \includegraphics[width=0.22\textwidth]{figures/smeared-images/before-component-removal-spw14} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-images/after-component-removal-spw14} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-images/rfi-like-component-spw14} &
        \includegraphics[width=0.22\textwidth]{figures/smeared-images/pickup-like-component-spw14} \\
        \textbf{(a)} & \textbf{(b)} & \textbf{(c)} & \textbf{(d)} \\
    \end{tabular}
    \caption{
        This figure illustrates the process of fitting for terrestrial sources of correlated noise.
        Each image is of the entire visible hemisphere above the OVRO-LWA but the data has been
        averaged over the entire 28 hour observing period. Because of this Cas A and Cyg A are
        smeared along tracks of constant declination. These images are generated using WSCLEAN
        \citep{2014MNRAS.444..606O}, uniform weighting, and baselines shorter than 15 wavelengths
        are flagged. \textbf{(a)} This image shows the initial data before any sources have been
        removed. \textbf{(b)} This image shows the final state after 3 terrestrial sources of
        correlated noise have been removed. \textbf{(c)} This image illustrates the contribution of
        an RFI source towards the north-west that was removed by peeling. \textbf{(d)} This removed
        component is not associated with a dot on the horizon.  Instead it is likely associated with
        common-mode pickup in the analog signal chain.
    }
    \label{fig:fitrfi}
\end{figure*}

When writing down equation~\ref{eq:basic-imaging}, it is implicitly assumed that the correlated
voltage fluctuations measured between pairs of antennas are exclusively generated by astronomical
sources of radio emission. In practice this implicit assumption can be violated. For instance a
low-frequency interferometer located in the vicinity of an arcing power line will see an additional
contribution from the radio-frequency interference (RFI) generated by the arcing process. Similarly
common-mode pickup along the analog signal path of the interferometer may generate an additional
spurious contribution to the measured visibilities. While the amplitude and phase of these
contaminating signals may fluctuate with time, they do not sweep across the sky at the sidereal rate
characteristic of astronomical sources.

The Owens Valley is an important source of water and power for the city of Los Angeles.
Unfortunately this means that high voltage power lines run along the valley to the west of the OVRO-
LWA. Some of these power line poles have faulty insulators that arc and produce pulsed, broadband
RFI. Because these poles exist in the near-field of the array, we have been able to localize some of
them by using the curvature of the incoming wavefront to infer a distance. Efforts are currently
underway to work with the utility pole owners to have these insulators replaced.

In the meantime it is possible to suppress their contamination in the dataset. The contribution of
these RFI sources to the visibilities can be plainly seen by averaging $>24$ hours of data with the
phase center set to zenith. In this way, true sky components are smeared along tracks of constant
declination while terrestrial sources (ie. the arcing power lines or any common-mode pickup) are not
smeared.  Obtaining a model for the RFI is complicated by the fact that the contaminating sources
are at extremely low elevations where the antenna response is essentially unknown (and inhomogeneous
due to antenna shadowing effects). Therefore it is not enough to know the physical location of the
faulty insulator generating the RFI. In addition you must also know the response of each antenna
(amplitude and phase) in the appropriate direction. This motivates the use of peeling
\citep{2008ISTSP...2..707M, 2015MNRAS.449.2668S}, which allows the antenna response to be a free
parameter.  Therefore model visibilities for the RFI can be obtained by peeling the sources after
smearing the visibilities over $>24$ hours. See Figure~\ref{fig:fitrfi} for an illustration of this
process.

However while attempting to peel RFI sources from the averaged visibilities, it was discovered that
frequently peeling would prefer to remove components from the visibilities that are not obviously
associated with any source on the horizon or elsewhere in the sky (see panel (d) in
Figure~\ref{fig:fitrfi}). In many cases the amplitude of these unassociated components exhibit the
same sawtooth oscillations indicative of gain fluctuations in the analog electronics even after
these gain fluctuations have been calibrated. This seems to imply that that these unassociated
components originate as common-mode pickup in the analog signal chain. That is, if the pickup occurs
somewhere in the middle of the analog signal chain, it will not see the same gain as the
astronomical signal. Therefore correcting the temperature dependence of the gains for the
astronomical signal does not correct the temperature dependence of the pick-up and hence it will
have a sawtooth pattern with respect to time.

@@

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = \b 0$.
In practice this is unlikely. In general the contamination is proportional to the overall amplitude
of the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/\|\b v_\text{terrestrial}\|$).

\todo{compute B*v for panels (c) and (d) to show that the pick-up is pretty bad. my expectation is
that it is actually worse than the RFI, but I don't actually know for sure}

These terrestrial sources do not rotate with the sky and hence their contamination tends to be
restricted to modes with small $m$. In this dataset the contamination is largely restricted to $m
\lesssim 3$. As a result if the contamination is not suppressed, it will manifest itself as rings
along stripes of constant declination. This effect is plainly visible in Figure \todo{make it}.

However because these rings are so distinctive in the final maps, it is possible to construct a
Wiener filter that removes them at the cost of losing a small amount of information about
astronomical sources.

\todo{describe how the Wiener filter works}

@@

A good model of the antenna beam is essential for widefield imaging. Because $m$-mode analysis
imaging operates on a full sidereal day of data, images are constructed after watching each point in
the sky move through large slices through the beam (excepting the celestial poles). The beam model
therefore serves two purposes:

\begin{enumerate}
    \item setting the flux scale as a function of declination
    \item reconciling observations from two separate sidereal times
\end{enumerate}

In the first case, all sources at a given declination take the same path through the antenna primary
beam. If the antenna response is overestimated along this track then all sources at this declination
will have their flux underestimated. Similarly if the antenna response is underestimated then all
the sources will have their flux overestimated.

\todo{discuss the possibility of declination dependent flux-scale errors in these maps}

The second case is more subtle. Sources are observed at a wide range of locations in the primary
beam of the antenna. The imaging process must reconcile all of these observations together. The beam
model essentially provides the instructions on how to do this. For example if at a time $t_1$ the
antenna gain towards a source is $g$ and at a later time $t_2$ the antenna gain towards that same
source (which has now moved) is $g/2$, then in order to correctly estimate the flux of the source
the observations from time $t_1$ need to be multiplied by $1/g$ but the observations from time $t_2$
need to be multiplied by $2/g$. A mistake in the beam model here will lead to a mistake in the
estimated flux.  However in general the amplitude and phase of the antenna response needs to be
known and so beam model errors can lead to observations being combined with incorrect phase as well.
This will lead to a degraded PSF in the final map. \todo{quick simulation of this?}

\subsection{The Ionosphere}

@@

This will tend to scatter power between $m$-modes. \todo{cite richard for this}

\todo{Simulate images of point sources with this}

\todo{I think the conclusion is sort of exactly what you would expect: refraction broadens the PSF
while scintillation effects the flux scale, but honestly not by much because we have such large
averages that unless the scintillation is Cauchy distributed it is not that bad..}

\subsection{Thermal Noise}

\todo{describe jack-knife tests}
